---
title: "Osteosarcoma scAtlas"
author: "Yogesh Budhathoki, Matt Cannon and Matt Gust"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
format:
    html:
        toc: true
        toc_float: true
        toc_depth: 5
        number_sections: false
        code-fold: true
        code-link: true
        df-print: kable
execute:
    cache: true
knitr:
    opts_chunk:
        cache.lazy: false
        cache.vars: ""
        warning: false
        message: false
lightbox: true
---

```{r setup}
#| include: FALSE
library(rrrSingleCellUtils)
library(Seurat)
library(ggrepel)
library(tidyverse)
library(harmony)
library(cowplot)
library(clustree)
library(data.table)
library(hdf5r)
library(Rmagic)
library(knitr)
library(scATOMIC)
library(copykat)
```

# Load in functions

{{< include child_qmds/helper_functions.qmd >}}

# Set up directory structure
```{bash mkdirs}
for directoryName in \
    output \
    output/figures \
    output/figures/gsea \
    output/figures/tumor_vs_stroma/clustree \
    output/counts \
    output/metadata \
    output/seurat_objects/sobj_preprocess \
    output/seurat_objects/comb_sobjs \
    output/seurat_objects/harmony_sobjs \
    output/seurat_objects/harmony_sobjs_annot \
    output/seurat_objects/sobj_preprocess_ \
    output/seurat_objects/tumor_vs_stroma \
    output/id_tumor/scATOMIC \
    output/id_tumor/SCEVAN \
    output/id_tumor/snvs/plots \
    output/id_tumor/celltype
do
    if [ ! -d ${directoryName} ]
    then
        mkdir -p ${directoryName}
    fi
done
```

# Read in raw data and process to Seurat objects
The metadata for all samples is stored in misc/all_samples.txt.

This reads in the raw data from the hdf5 files and processes them to Seurat objects.

The individual objects are saved to `output/seurat_objects/sobj_preprocess/`

Merged data for each group (as defined in the metadata) are saved to `output/seurat_objects/comb_sobjs/`

Harmonized datasets are saved to `output/seurat_objects/sobj_preprocess_merged/`

{{< include child_qmds/read_raw_data.qmd >}}

# Use SingleR and reference datasets to annotate cell types
The output of this is written to `output/seurat_objects/harmony_sobjs_annot/`

{{< include child_qmds/annotate_celltypes.qmd >}}

# Use SCEVAN and scATOMIC to try to identify tumor cells
This will help in splitting the cancer cells from the stroma cells.

{{< include child_qmds/annot_cancer_scevan_scatomic.qmd >}}

# Use SingleR cell annotations to try to ID cancer cells

{{< include child_qmds/annot_cancer_celltype.qmd >}}

# Split cancer vs the stroma

{{< include child_qmds/split_tumor_stromal.qmd >}}

# Run differential expression and GSEA

{{< include child_qmds/degs_gseaplots.qmd >}}



# Session info
```{r}
sessionInfo()
```
