---
title: "Osteosarcoma scAtlas"
output: 
    html_document:
        code_folding: hide
always_allow_html: true
---

# Load the library
```{r, include=FALSE}
library(rrrSingleCellUtils)
library(Seurat)
library(ggrepel)
library(tidyverse)
library(stringr)
library(harmony)
library(cowplot)
library(clustree)
library(data.table)
library(hdf5r)
library(Rmagic)
library(scATOMIC)
library(SCEVAN)
library(knitr)
```

# load in functions as a child document
```{r, child='function_atlas.Rmd'}
```

# setup
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      cache.lazy = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

## Read metadata
This metadata is has information on sample and all the details on methods used,
source, organism, unique identifier per group, QC, and other relevant information
```{r, cache.vars="all_samples_csv"}
all_samples_csv <- 
    read_tsv("allsample_details.txt")

all_samples_csv$sp_pattern <-
    str_replace_all(all_samples_csv$sp_pattern, "nothing", "")

# #for the sake of presentation, I will just have patient data
# all_samples_csv <- filter(all_samples_csv,
#                           unique == "patient_prim" | unique =="patient_mets")
```

# Count before the QC
Count the number of cells per sample before the QC
```{r, fig.width = 10, fig.height = 10, fig.align = "center", echo = FALSE}
single_sobjs <- 
    list.files(path = "output/seurat_objects/sobj_preprocess/",
               full.names = FALSE)
counts <- tibble()
for (item in single_sobjs){
    sobj <- qs::qread(str_c("output/seurat_objects/sobj_preprocess/",
                            item))
    sobj$ncount_median_before <- median(sobj$nCount_RNA)
    unique <- sobj@meta.data$unique %>%
        head(n=1)
    Hmeta_data <- sobj@meta.data %>%
        as.data.table %>%
        group_by(sample_name, unique) %>%
        summarize(ncount_median_before = median(nCount_RNA),
                  N_cells_before = n(),
                  .groups = "drop")

    counts <- Hmeta_data %>%
        select("sample_name", "unique", "ncount_median_before", "N_cells_before") %>%
        rbind(counts) %>% 
        arrange(unique)
}
```

# Count post QC
some of the samples are discarded as the ncountmin was set 500 and number of 
cells limit was set to 200 cells and above per sample. The upper cutoff was set 
as 5000 max cells per sample for similar dataset representation 
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
counts_afterQC <- tibble()

for (group in unique(all_samples_csv$unique)) {
    object <- qs::qread(str_c("output/seurat_objects/harmony_sobjs/",
                              group,
                              ".qs"))
    #DimPlot(object)
    #extract meta data
    Hmeta_data <- object@meta.data %>%
        as.data.table %>%
        group_by(sample_name, unique) %>% 
        summarise(ncount_median_after = median(nCount_RNA),
                  N_cells_after = n(),
                  .groups = "drop")

    counts_afterQC <- Hmeta_data %>% 
        select("N_cells_after", "sample_name", "unique", "ncount_median_after") %>%
        rbind(counts_afterQC) %>%
        arrange(unique)
}

combined_table <- 
    dplyr::full_join(counts, counts_afterQC, by = c("sample_name", "unique")) %>%
    select("sample_name", "unique", "ncount_median_before", "ncount_median_after", "N_cells_before", "N_cells_after")

DT::datatable(arrange(combined_table, unique, desc(N_cells_before)))
```


# Data Integration DimPlots {.tabset}
This is the dimplot of the integrated data. The data is integrated using harmony.
The plots will be in order of before and after integration for each group in
tabset. The plots are grouped by sample name, method, and data source
```{r, fig.width = 10, fig.height = 10, fig.align = "center", echo = FALSE}
BI_clusters <- list()
BI_method <- list()
BI_data_source <- list()

AI_clusters <- list()
AI_method <- list()
AI_data_source <- list()

plot_my_stuff <- function(object, what, ncol = 1) {
    DimPlot(object,
            group.by = what,
            ncol = ncol,
            label = T,
            repel = T,
            shuffle = T,
            cols = c(plot_cols, sample(rainbow(1000)))) +
        coord_fixed()
}

for (group in unique(all_samples_csv$unique)){
    before_harmo_obj <- qs::qread(str_c("output/seurat_objects/comb_sobjs/",
                                        group,
                                        ".qs"))
    BI_clusters[[group]] <- plot_my_stuff(before_harmo_obj, "sample_name")
    BI_method[[group]] <- plot_my_stuff(before_harmo_obj, "method")
    BI_data_source[[group]] <- plot_my_stuff(before_harmo_obj, "data_source")

    harmo_obj <- qs::qread(str_c("output/seurat_objects/harmony_sobjs/",
                                group,
                                ".qs"))
    #optimize resolution, silhouette score does not work for large dataset
    min_cell_silhouette <- 65000
    n_cells <- length(Cells(harmo_obj))
    keep_n_cells <- min(min_cell_silhouette, n_cells)
    set.seed(63249528)
    res_score <-
        optimize_silhouette(sobject = harmo_obj[, sample(1:n_cells, keep_n_cells)],
                            test_res = seq(0.1, 0.9, by = 0.1),
                            summary_plot = FALSE,
                            reduction = "harmony") %>%
        filter(sil_vals == max(sil_vals)) %>%
        slice_head(n=1) %>%
        dplyr::pull(res_vals)
    print(str_c("Optimized resolution for ", group, " is ", res_score))

    AI_clusters[[group]] <- plot_my_stuff(harmo_obj, what = "sample_name")
    AI_method[[group]] <- plot_my_stuff(harmo_obj, "method")
    AI_data_source[[group]] <- plot_my_stuff(harmo_obj, "data_source")

}
```

## Plots for patient_prim before and after harmony integration
### Before Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
BI_clusters$patient_prim
BI_method$patient_prim
BI_data_source$patient_prim
```

### After Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
AI_clusters$patient_prim
AI_method$patient_prim
AI_data_source$patient_prim
```

## Plots for patient_mets before and after harmony integration
### Before Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
BI_clusters$patient_mets
BI_method$patient_mets
BI_data_source$patient_mets
```

### After Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
AI_clusters$patient_mets
AI_method$patient_mets
AI_data_source$patient_mets
```

## Plots for Xeno_prim_human before and after harmony integration
### Before Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
BI_clusters$xeno_prim_human
BI_method$xeno_prim_human
BI_data_source$xeno_prim_human
```

### After Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
AI_clusters$xeno_prim_human
AI_method$xeno_prim_human
AI_data_source$xeno_prim_human
```

## Plots for Xeno_prim_mouse before and after harmony integration
### Before Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
BI_clusters$xeno_prim_mouse
BI_method$xeno_prim_mouse
BI_data_source$xeno_prim_mouse
```

### After Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
AI_clusters$xeno_prim_mouse
AI_method$xeno_prim_mouse
AI_data_source$xeno_prim_mouse
```

## Plots for xeno_mets_human before and after harmony integration
### Before Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
BI_clusters$xeno_mets_human
BI_method$xeno_mets_human
BI_data_source$xeno_mets_human
```

### After Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
AI_clusters$xeno_mets_human
AI_method$xeno_mets_human
AI_data_source$xeno_mets_human
```

## Plots for xeno_mets_mouse before and after harmony integration
### Before Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
BI_clusters$xeno_mets_mouse
BI_method$xeno_mets_mouse
BI_data_source$xeno_mets_mouse
```

### After Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
AI_clusters$xeno_mets_mouse
AI_method$xeno_mets_mouse
AI_data_source$xeno_mets_mouse
```

## Plots for mm_prim before and after harmony integration
### Before Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
BI_clusters$mm_prim 
BI_method$mm_prim 
BI_data_source$mm_prim 
```

### After Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
AI_clusters$mm_prim 
AI_method$mm_prim 
AI_data_source$mm_prim 
```

## Plots for mm_mets  before and after harmony integration
### Before Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
BI_clusters$mm_mets  
BI_method$mm_mets  
BI_data_source$mm_mets 
```

### After Integration
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
AI_clusters$mm_mets  
AI_method$mm_mets  
AI_data_source$mm_mets  
```

# Clustering and setting the right resolution {.tabset}
Used silhouette score and clustree method for right resolution. For this step, 
silhouette score was used for clustering. The silhouette score is a measure of
how similar an object is to its own cluster (cohesion) compared to other clusters
(separation). The silhouette ranges from -1 to 1, where a high value indicates
that the object is well matched to its own cluster and poorly matched to
neighboring clusters. If most objects have a high value, then the clustering
solution is appropriate. If many points have a low or negative value, then the
clustering solution may have too many or too few clusters. The silhouette score
does not work for large dataset (cells above around 65000), so the clustree method
was used to compliment the right resolution. 
```{r, fig.width = 10, fig.height = 10, fig.align = "center", echo = FALSE, message = FALSE, warning=FALSE}
clustrees <- list()
silhouette_scores <- list()

for (group in unique(all_samples_csv$unique)) {
    object <- qs::qread(str_c("output/seurat_objects/harmony_sobjs/",
                              group,
                              ".qs"))

    resolution_range <- seq(from = 0, to = 1, by = 0.05)
    clustree_sobj <- FindClusters(object,
                                  resolution = resolution_range)
    tree <- clustree::clustree(clustree_sobj,
                               prefix = "RNA_snn_res.")

    clustrees[[group]] <- tree

    #optimize resolution, silhouette score does not work for large dataset
    min_cell_silhouette <- 65000
    n_cells <- length(Cells(object))
    keep_n_cells <- min(min_cell_silhouette, n_cells)
    set.seed(63249528)
    res_score <-
        optimize_silhouette(sobject = object[, sample(1:n_cells, keep_n_cells)],
                            test_res = seq(0, 1, by = 0.05),
                            summary_plot = FALSE,
                            reduction = "harmony") %>%
        filter(sil_vals == max(sil_vals)) %>%
        slice_head(n=1) %>%
        dplyr::pull(res_vals)
    print(str_c("Optimized resolution for ", group, " is ", res_score))

    silhouette_scores[[group]] <- res_score
}

```

## Clustering patient prim
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustrees$patient_prim
silhouette_scores$patient_mets
```

## Clustering patient mets
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustrees$patient_mets
silhouette_scores$patient_mets
```

## Clustering xenograft primary human
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustrees$xeno_prim_human
silhouette_scores$xeno_prim_human
```

## Clustering xenograft primary mouse
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustrees$xeno_prim_mouse
silhouette_scores$xeno_prim_mouse
```

## Clustering xenograft mets human
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustrees$xeno_mets_human
silhouette_scores$xeno_mets_human
```

## Clustering xenograft mets mouse
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustrees$xeno_mets_mouse
silhouette_scores$xeno_mets_mouse
```

## Clustering mouse primary
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustrees$mm_prim
silhouette_scores$mm_prim
```

## Clustering mouse mets
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustrees$mm_mets
silhouette_scores$mm_mets
```

# Make plots for annotated datasets {.tabset}
Each seurat obejct was annotated using the `annotate` function. The function uses
the `SingleR` package to annotate the cells. The `SingleR` package uses a 
reference dataset to annotate the cells. The reference dataset used here is the
`HumanPrimaryCellAtlas` and `MousePrimaryCellAtlas` for human and mouse samples
respectively. I also input some manually annotated dataset to the reference such
as human_lung, human_bone, mouselung.
Annotation was by one level of Recurlstering. The Recurlstering was done using
running proess seurat and harmony integration. Each cluster was then added back
by the `AddMetaData` function.
```{r annotated, fig.width = 10, fig.height = 10, fig.align = "center", echo = FALSE}
singler_annotation <- list()
anotation_cluster_assignment <- list()
feature_plots <- list()

for (group in unique(all_samples_csv$unique)) {
    object <- qs::qread(str_c("output/seurat_objects/harmony_sobjs/",
                              group,
                              ".qs"))
    singler_annotation[[group]] <- plot_my_stuff(object, what = "annotations")
    anotation_cluster_assignment[[group]] <- plot_my_stuff(object, "new_annot_clust")

    if (object$organism[1] == "human") {
        feature_plots[[group]] <- FeaturePlot(object,
                                              features = c("COL1A1", "COL1A2", "SATB2", "RUNX2"))
    } else {
        feature_plots[[group]] <- FeaturePlot(object,
                                              features = c("Col1a1", "Col1a2", "Satb2", "Runx2"))
    }
}
```

## Patient prim 
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
singler_annotation$patient_prim
anotation_cluster_assignment$patient_prim
feature_plots$patient_prim
```

## Patient mets
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
singler_annotation$patient_mets
anotation_cluster_assignment$patient_mets
feature_plots$patient_mets
```

## Xenograft primary human
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
singler_annotation$xeno_prim_human
anotation_cluster_assignment$xeno_prim_human
feature_plots$xeno_prim_human
```

## Xenograft primary mouse
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
singler_annotation$xeno_prim_mouse
anotation_cluster_assignment$xeno_prim_mouse
feature_plots$xeno_prim_mouse
```

## Xenograft mets human
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
singler_annotation$xeno_mets_human
anotation_cluster_assignment$xeno_mets_human
feature_plots$xeno_mets_human
```

## Xenograft mets mouse
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
singler_annotation$xeno_mets_mouse
anotation_cluster_assignment$xeno_mets_mouse
feature_plots$xeno_mets_mouse
```

## Mouse primary
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
singler_annotation$mm_prim
anotation_cluster_assignment$mm_prim
feature_plots$mm_prim
```

## Mouse mets
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
singler_annotation$mm_mets
anotation_cluster_assignment$mm_mets
feature_plots$mm_mets
```

# SCEVAN and scATOMIC plots {.tabset}
The SCEVAN and scATOMIC plots are used to visualize the cell type assignment as
found by the algorithms. The plots are grouped by the sample name and the
cell type assignment. The plots are also grouped by the cancer vs normal cells.
scAtomic works only with human datasets and SCEVAN works with both human and mouse.
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
scevan_plots <- list()
scATOMIC_plots <- list()
for (group in c("patient_prim", "patient_mets", "mm_prim", "mm_mets")) {
    scevan_object <- qs::qread(str_c("output/SCEVAN/",
                                    group,
                                    ".qs"))
    scevan_plots[[group]] <- plot_my_stuff(scevan_object, "class")
}

for (group in c("patient_prim", "patient_mets")) {
    scATOMIC_object <- qs::qread(str_c("output/scATOMIC/",
                                    group,
                                    ".qs"))
    scATOMIC_plots[[group]] <- plot_my_stuff(scATOMIC_object, "pan_cancer_cluster")
}
```

## Patient primary
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
scevan_plots$patient_prim
scATOMIC_plots$patient_prim
```

## Patient mets
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
scevan_plots$patient_mets
scATOMIC_plots$patient_mets
```

## Mouse primary
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
scevan_plots$mm_prim
```

## Mouse mets
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
scevan_plots$mm_mets
```

# silhouette score and clustree {.tabset}
The results from the annotation, featureplots and the SCEVAN and scATOMIC plots
are used to distinguish the cancer vs normal cells. Once the cells are split into
cancer and normal cells, the silhouette score and clustree method are used to 
find the right resolution for the clustering. 
```{r, fig.width = 10, fig.height = 10, fig.align = "center", echo = FALSE}
object_list <-
    c("patient_prim_cancer_cells",
    "patient_mets_cancer_cells",
    "mm_prim_cancer_cells",
    "mm_mets_cancer_cells",
    "xeno_prim_human",
    "xeno_mets_human",
    "patient_prim_normal_cells",
    "patient_mets_normal_cells",
    "xeno_prim_mouse",
    "xeno_mets_mouse",
    "mm_prim_normal_cells",
    "mm_mets_normal_cells")

# Calculate cell type percentage for each object
clustree_list <- list()
silhouette_list <- list()

for (group in object_list) {
    s_obj <- 
        qs::qread(str_c("output/seurat_objects/tumor_vs_stroma/",
                              group,
                              ".qs"))
    
    #optimize resolution, silhouette score does not work for large dataset
    min_cell_silhouette <- 65000
    n_cells <- length(Cells(s_obj))
    keep_n_cells <- min(min_cell_silhouette, n_cells)

    set.seed(1111471)
    res_normal <-
        optimize_silhouette(sobject = s_obj[, sample(1:n_cells, keep_n_cells)],
                            test_res = seq(0.05, 0.9, by = 0.05),
                            summary_plot = FALSE,
                            reduction = "harmony") %>%
        arrange(desc(sil_vals), desc(res_vals)) %>%
        filter(sil_vals == max(sil_vals)) %>%
        slice_head(n=1) %>%
        dplyr::pull(res_vals)

    silhouette_list[[group]] <- res_normal

    resolution_range <- seq(from = 0, to = 1, by = 0.05)
    clustree_sobj <- FindClusters(s_obj,
                                    resolution = resolution_range)
    tree <- clustree::clustree(clustree_sobj,
                                prefix = "RNA_snn_res.") + 
        ggtitle(str_c("Clustree for ", group))
    clustree_list[[group]] <- tree
}
```

## Patient primary cancer cells
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustree_list$patient_prim_cancer_cells
silhouette_list$patient_prim_cancer_cells
```

## Patient mets cancer cells
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustree_list$patient_mets_cancer_cells
silhouette_list$patient_mets_cancer_cells
```

## Mouse primary cancer cells
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustree_list$mm_prim_cancer_cells
silhouette_list$mm_prim_cancer_cells
```

## Mouse mets cancer cells
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustree_list$mm_mets_cancer_cells
silhouette_list$mm_mets_cancer_cells
```

## Xenograft primary human
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustree_list$xeno_prim_human
silhouette_list$xeno_prim_human
```

## Xenograft mets human
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustree_list$xeno_mets_human
silhouette_list$xeno_mets_human
```

## Patient primary normal cells
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustree_list$patient_prim_normal_cells
silhouette_list$patient_prim_normal_cells
```

## Patient mets normal cells
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustree_list$patient_mets_normal_cells
silhouette_list$patient_mets_normal_cells
```

## Xenograft primary mouse
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustree_list$xeno_prim_mouse
silhouette_list$xeno_prim_mouse
```

## Xenograft mets mouse
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustree_list$xeno_mets_mouse
silhouette_list$xeno_mets_mouse
```

## Mouse primary normal cells
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustree_list$mm_prim_normal_cells
silhouette_list$mm_prim_normal_cells
```

## Mouse mets normal cells
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
clustree_list$mm_mets_normal_cells
silhouette_list$mm_mets_normal_cells
```

# UMAPS combined and individual {.tabset}
The UMAPS are plotted for the combined and individual datasets. The combined
umaps have both cancer and stromal cells together in the same plot. The individual
umaps have the cancer and stromal cells in separate plots. The umaps are plotted
for the patient primary, patient mets, xenograft primary, xenograft mets, mouse
primary, and mouse mets. 
```{r plotsumaps, fig.width = 10, fig.height = 10, fig.align = "center", echo = FALSE}
#Checkups to make sure the outputs make sense

umaps_group <-
    tribble(~sobj,            ~title,                  ~combine,
            "patient_prim",    "Patient Primary",      "no",
            "patient_mets",    "Patient Metastatic",   "no",
            "xeno_prim",       "Xenograft Primary",    "yes",
            "xeno_mets",       "Xenograft Metastatic", "yes",
            "mm_prim",         "Mouse Primary",        "no",
            "mm_mets",         "Mouse Metastatic",     "no")

annotated <- list()
merged <- list()
just_tumor <- list()
just_stroma <- list()

for (item in seq_len(nrow(umaps_group))) {
    sobj_name <- umaps_group$sobj[item]
    combine <- umaps_group$combine[item]
    title <- umaps_group$title[item]
    if (combine == "no") {
        object <-
            qs::qread(str_c("output/seurat_objects/annotated_sobjs/",
                            sobj_name,
                            ".qs"))
        annotated[[sobj_name]] <- 
            plot_my_stuff(object,
                          what = c("new_annot_clust", "seurat_clusters")) +
                ggtitle("Annotated cell types by cluster")
        merged[[sobj_name]] <- 
            plot_my_stuff(object,
                          what = c("cell_group", "sample_name")) +
                ggtitle("Tumor and Stromal cells")

        #plot the tumor cells
        tumor <-
            qs::qread(str_c("output/seurat_objects/resolution_optimized/",
                            sobj_name,
                            "_cancer_cells.qs"))
        just_tumor[[sobj_name]] <- 
            plot_my_stuff(tumor,
                          what = c("sample_name", "new_annot_clust", "seurat_clusters")) +
                ggtitle("Tumor cells reclustering")

        #plot the stromal cells
        stroma <- 
            qs::qread(str_c("output/seurat_objects/resolution_optimized/",
                            sobj_name,
                            "_normal_cells.qs"))
        just_stroma[[sobj_name]] <- 
            plot_my_stuff(stroma,
                          what = c("sample_name", "seurat_clusters", "new_annot_clust")) +
                ggtitle("Stromal cells reclustering")
    } else {
        first_obj <- 
            qs::qread(str_c("output/seurat_objects/resolution_optimized/",
                            sobj_name,
                            "_human.qs"))
        just_tumor[[sobj_name]] <-
            plot_my_stuff(first_obj,
                          what = c("sample_name", "seurat_clusters", "new_annot_clust")) +
                ggtitle("Tumor cells")

        second_obj <- 
            qs::qread(str_c("output/seurat_objects/resolution_optimized/",
                            sobj_name,
                            "_mouse.qs"))
        just_stroma[[sobj_name]] <- 
            plot_my_stuff(second_obj,
                          what = c("sample_name", "seurat_clusters", "new_annot_clust")) +
                ggtitle("Stromal cells")
    }
}

```

## UMAPS Patient primary
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
annotated$patient_prim
merged$patient_prim
just_tumor$patient_prim
just_stroma$patient_prim
```

## UMAPS Patient mets
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
annotated$patient_mets
merged$patient_mets
just_tumor$patient_mets
just_stroma$patient_mets
```

## UMAPS Xenograft primary
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
just_tumor$xeno_prim
just_stroma$xeno_prim
```

## UMAPS Xenograft mets
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
just_tumor$xeno_mets
just_stroma$xeno_mets
```

## UMAPS Mouse primary
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
annotated$mm_prim
merged$mm_prim
just_tumor$mm_prim
just_stroma$mm_prim
```

## UMAPS Mouse mets
```{r, fig.width = 10, fig.height = 10, fig.align = "center"}
annotated$mm_mets
merged$mm_mets
just_tumor$mm_mets
just_stroma$mm_mets
```

# calculate the celltype percentage and score per each clusters
The cell type percentage and score per each cluster is calculated. The cell type
percentage is calculated by taking the top 5 cell types and their percentage
per cluster. The cell type score is calculated by taking the median of the cell
score per cluster. The cell type percentage and score is calculated for the
patient primary cancer cells, patient mets cancer cells, mouse primary cancer
cells, mouse mets cancer cells, xenograft primary human, xenograft mets human,
patient primary normal cells, patient mets normal cells, xenograft primary mouse,
xenograft mets mouse, mouse primary normal cells, and mouse mets normal cells.
```{r, fig.width = 10, fig.height = 10, fig.align = "center", echo = FALSE}
object_list <- 
    c("patient_prim_cancer_cells",
    "patient_mets_cancer_cells",
    "mm_prim_cancer_cells",
    "mm_mets_cancer_cells",
    "xeno_prim_human",
    "xeno_mets_human",
    "patient_prim_normal_cells",
    "patient_mets_normal_cells",
    "xeno_prim_mouse",
    "xeno_mets_mouse",
    "mm_prim_normal_cells",
    "mm_mets_normal_cells")

# Calculate cell type percentage for each object
all_data_celltypes <- tibble()
for (group in object_list) {
    object <- 
        qs::qread(str_c("output/seurat_objects/resolution_optimized/",
                              group,
                              ".qs"))
    # plot_my_stuff(object, what = c("seurat_clusters", "new_annot_clust", "sample_name")) +
    #     ggtitle(group)
    object$cell_types_string <- "NA"
    object$med_cellscore_string <- "NA"
    celltype_df <-
        object@meta.data %>%
        select(seurat_clusters) %>%
        as_tibble() %>%
        mutate(group = group,
               assigned_celltype = NA,
               cell_types_string = NA,
               med_cellscore_string = NA,
               NCH_assignment = NA,
               NCI_assignment = NA,
               SJ_assignment = NA) %>%
        distinct() %>%
        arrange(seurat_clusters)
    for (i in seq_len(nrow(celltype_df))) {
        temp_obj <-
            subset(x = object,
                   seurat_clusters == celltype_df$seurat_clusters[i])
        cell_types_percent <- 
            table(temp_obj$annotations) %>%
            as.data.frame() %>%
            arrange(desc(Freq), .by_group = TRUE) %>%
            slice_head(n=5) %>%
            mutate(Percentage = Freq / sum(Freq) * 100) %>%
            group_by(Var1) %>%
            mutate(med_cellscore = median(temp_obj$cell_scores[temp_obj$annotations == Var1]) %>% 
                round(2))

        celltype_df$assigned_celltype[i] <- 
            cell_types_percent$Var1[1] %>%
            as.character()

        # Create a string of Var1 and Percentage
        celltype_df$cell_types_string[i] <- 
            paste(cell_types_percent$Var1,
                  cell_types_percent$Percentage %>% round(1),
                  sep = ":",
                  collapse = ",")

        celltype_df$med_cellscore_string[i] <- 
            paste(cell_types_percent$Var1, 
                  cell_types_percent$med_cellscore,
                  sep = ":",
                  collapse = ",")
    }

    all_data_celltypes <-
        celltype_df %>%
        rbind(all_data_celltypes)

    # # Append the strings to the main object
    # # for (clusters in unique(object$seurat_clusters)) {
    # #     object$cell_types_string[object$seurat_clusters == clusters] <- celltype_df[[`clusters`]]$cell_types_string
    # #     object$med_cellscore_string[object$seurat_clusters == clusters] <- celltype_df[[`clusters`]]$med_cellscore_string
    # # }
    # qs::qsave(object, 
    #           file = str_c("output/seurat_objects/tumor_vs_stroma/",
    #                                    group,
    #                                    ".qs"))
}

write_tsv(all_data_celltypes, "output/all_data_celltypes.tsv")
```

# load in GSEA childdocs
```{r, child='degs_gseaplots.rmd'}
```