---
title: "Human Osteosarcoma Primary - Clustering"
subtitle: "Cell type annotation reference assembly"
output: html_document
---

This script begins with the output from Primary-CreateData (comb.qs), merges and
integrates the samples, performs clustering, initial identification of cell subtypes,
focused on tumor cells vs stromal cells, and returns a file identifying the tumor
cells (selection.rds).

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  dev = "png",
  fig.width = 6,
  fig.height = 6,
  fig.asp = NULL
)

library(rrrSingleCellUtils)
library(Seurat)
library(ggplot2)
library(ggrepel)
library(tidyverse)
library(stringr)
library(harmony)
library(cowplot)

set.seed(888)

sysfonts::font_add("Arial Narrow", "ARIALN.TTF", bold = "ARIALNB.TTF")
showtext::showtext_auto()

theme_set(ggpubr::theme_pubr())

plot_cols <- c("#D43F3AFF", "#EEA236FF", "#357EBDFF", "#5CB85CFF", "#B8B8B8FF",
  "#9632B8FF",  "#46B8DAFF", "#90302DFF", "#A66D04FF", "#2D577FFF", "#3E7E3EFF",
  "#7D7D7DFF", "#6D1D87FF", "#097F9AFF", "#FF6E6AFF", "#FFBB70FF", "#68A4E3FF",
  "#79D379FF", "#CDCDCDFF", "#BF6BE2FF", "#69D1F3FF")

rDimPlot <- function(object, title = "Unlabeled Plot",
  label = T, pt.size = 1, ...) {
  if(length(levels(Idents(object))) < 22) {  
    p <- Seurat::DimPlot(
      object = object, 
      label = label, 
      pt.size = pt.size,
      cols = alpha(plot_cols, 0.6),
      ...) +
      ggtitle(title) +
      theme(legend.position = "none") +
      coord_fixed()
    print(p)
    return(p)
  } else {
    print("Too many identity classes to use this function. Requires <22.")
  }
}

rFeaturePlot <- function(object, features, title, pt.size = 1, order = T, ...) {
  p <- Seurat::FeaturePlot(
    object = object,
    features = features,
    pt.size = pt.size,
    order = order,
    cols = (c("lightgoldenrod", "darkred")),
    ...) +
    ggtitle(title) +
    coord_fixed()
  print(p)
  return(p)
}
```

# Reload the dataset

This script picks up where the "Primary-CreateData.Rmd" script left off, creating (and
here reloading) the "comb.qs" object.

```{r}
# Start from this stopping point - merged and harmony aligned
comb <- qs::qread("HuOsteo/comb.qs")

# Plot the integrated samples
rDimPlot(comb, "Primary Tumors (integrated)", repel = T)

rDimPlot(comb, "Primary Tumors (integrated)", group.by = "src")

rDimPlot(comb, "Primary Tumors (integrated)", split.by = "src", ncol = 4)
```

## Interim conclusions

A broad array of distinct cell types is represented across the different primary
tumor samples. These subsets appear to be of high enough quality and are appropriate
for building both an tumor cell and a stromal reference.

# A brief tangent - Optimize UMAP parameters

This code was created to illustrate how different changes in UMAP parameters affect
the way that cells cluster. Bottom line is that the default parameters work pretty
well and are fairly balanced, though tweaking can be helpful in certain situations.

```{r}
# p <- list()
# for(i in 1:5) {
#   test <- RunUMAP(comb, reduction = "harmony", dims = 1:30,
#     set.op.mix.ratio = (i-1)*0.25)
#   test <- test %>%
#     FindNeighbors(reduction = "harmony", dims = 1:30) %>%
#     FindClusters()
#   p[[i]] <- rDimPlot(test, reduction = "umap", label = T, repel = T) +
#       coord_fixed() +
#       ggtitle(str_c("set.op.mix.ratio = ", (i-1)*0.25)) +
#       theme(legend.position = "none")
# }
# plot_grid(plotlist = p, ncol = 3)

# p <- list()
# for(i in 1:5) {
#   test <- RunUMAP(comb, reduction = "harmony", dims = 1:30, 
#     n.neighbors = (i * 10))
#   test <- test %>%
#     FindNeighbors(reduction = "harmony", dims = 1:30) %>%
#     FindClusters()
#   p[[i]] <- rDimPlot(test, reduction = "umap", label = T, repel = T) +
#       coord_fixed() +
#       ggtitle(str_c("n.neighbors = ", (i * 10))) +
#       theme(legend.position = "none")
# }
# plot_grid(plotlist = p, ncol = 3)

# p <- list()
# for(i in 1:5) {
#   test <- RunUMAP(comb, reduction = "harmony", dims = 1:30, 
#     n.epochs = (i * 100))
#   test <- test %>%
#     FindNeighbors(reduction = "harmony", dims = 1:30) %>%
#     FindClusters()
#   p[[i]] <- rDimPlot(test, reduction = "umap", label = T, repel = T) +
#     coord_fixed() +
#     ggtitle(str_c("n.epochs = ", (i * 100))) +
#     theme(legend.position = "none")
# }
# plot_grid(plotlist = p, ncol = 3)

# p <- list()
# for(i in 1:5) {
#   test <- RunUMAP(comb, reduction = "harmony", dims = 1:30, 
#     min.dist = (.00045 * 4 ^ i))
#   test <- test %>%
#     FindNeighbors(reduction = "harmony", dims = 1:30) %>%
#     FindClusters()
#   p[[i]] <- rDimPlot(test, reduction = "umap", label = T, repel = T) +
#     coord_fixed() +
#     ggtitle(str_c("min.dist = ", (.00045 * 4 ^ i))) +
#     theme(legend.position = "none")
# }
# plot_grid(plotlist = p, ncol = 3)

# p <- list()
# for(i in 1:5) {
#   test <- RunUMAP(comb, reduction = "harmony", dims = 1:30, 
#     local.connectivity = (2 ^ i))
#   test <- test %>%
#     FindNeighbors(reduction = "harmony", dims = 1:30) %>%
#     FindClusters()
#   p[[i]] <- rDimPlot(test, reduction = "umap", label = T, repel = T) +
#     coord_fixed() +
#     ggtitle(str_c("local.connectivity = ", (2 ^ i))) +
#     theme(legend.position = "none")
# }
# plot_grid(plotlist = p, ncol = 3)
```

# Evaluate for cell cycle

```{r}
comb <- kill_cc(comb)
```

# Begin cell type annotation

Begin the process of identifying the cell types and annotating the samples. We will first
need to identify the tumor cells and separate them away from the stromal cells so that we
can focus in on the intratumoral heterogeneity.

First, we will use some standard assays to look for canonical markers of osteosarcoma cells.

```{r}
feats <- c("COL1A1", "COL1A2", "SATB2")

for(f in feats) {
  p <- rFeaturePlot(comb, f, features = f)
  print(p)
}
```

Next, we will look at the markers that were used in the manuscript that reported the
GSE152048 dataset. I'm not sure that these are the ideal markers, but they are somewhat
useful, and this allows comparison to the published record.

```{r, fig.height = 8, fig.width = 8}
ms <- list(
  Osteoblastic = c("RUNX2", "COL1A1", "CDH11", "IBSP"),
  Chondroblastic = c("SOX9", "ACAN", "PTH1R"),
  Osteoclast = c("ACP5", "CTSK", "MMP9"),
  Myeloid = c("CD74", "CD14", "FCGR3A"),
  TCell = c("CD3E", "IL7R", "CD8A", "CD4", "NKG7"),
  NKCell = c("NKG7", "GNLY"),
  NKTCell = c("NKG7", "GNLY", "CD3E"),
  DCCell = c("CD1C", "FCER1A", "CLEC9A", "CCR7", "CD14", "CD163"),
  Fibroblast = c("DCN", "COL1A1"),
  Pericyte = c("RGS5", "ACTA2"),
  MSC = c("MME", "THY1", "CXCL12", "SFRP2"),
  Endothelial = c("PECAM1", "VWF"),
  Myoblast = c("MYL1", "MYLPF"),
  BCell = c("MS4A1", "CD19", "JCHAIN")
)

g <- list()
for(f in names(ms)) {
  comb <- AddModuleScore(comb, ms[[f]], name = f)
  g[[f]] <- rFeaturePlot(comb, f, features = str_c(f, "1"))
}

plot_grid(
  plotlist = g,
  labels = LETTERS[seq_len(length(g))],
  ncol = 4, nrow = 4
)
```

# Isolate and subcluster the tumor cells

```{r}
if (!file.exists("selection.rds")) {
  selection <- CellSelector(p)
  saveRDS(selection, "selection.rds")
}
selection <- readRDS("selection.rds")
Idents(comb) <- comb$seurat_clusters
Idents(comb, cells = selection) <- "Tumor"

rDimPlot(comb, "Isolated (likely) tumor cells")
```

Focus in on a module that defines the tumor cells so that we can isolate them.
Given the purpose in creating an osteosarcoma signature, we will use an ROC
statistic to create the DEG list.

```{r}
tmarks_roc <- FindMarkers(comb, ident.1 = "Tumor",
  test.use = "roc")
tmarks_roc_labs <- tmarks_roc[
  tmarks_roc$power > .75 &
  abs(tmarks_roc$avg_log2FC) > 2.2, ]

tmarks_roc_labs$gene <- rownames(tmarks_roc_labs)
ggplot(tmarks_roc, aes(avg_log2FC, power)) +
  geom_point() +
  geom_label_repel(data = tmarks_roc_labs,
    aes(label = gene), max.overlaps = 20) +
  ggtitle("Genes that identify tumor cells")

os_sig <- rownames(tmarks_roc_labs[
  tmarks_roc_labs$avg_log2FC > 0, ])
comb <- AddModuleScore(comb, os_sig, name = "os_score")
rFeaturePlot(comb, "OS Signature Score", features = "os_score1",
  min.cutoff = 1, max.cutoff = 5)
```

After several different criteria strict vs loose for defining the genes
that should comprise the osteosarcoma signature, it appears that the simplest
gene list based off of the strictest power/log2FC criteria performs the best.

The "selection.rds" file generated here together with the "comb.qs" file generated
in the "Primary-CreateData.Rmd" script will be used to isolate tumor and stromal
cells in the "Primary-AnnotateTumor.Rmd" and the "Primary-AnnotateStroma.Rmd" scripts.