---
title: "Human Osteosarcoma Primary - Annotate Tumor Cells"
subtitle: "Cell type annotation reference assembly"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE,
  dev="svg",
  fig.width=4,
  fig.height=4,
  fig.asp=NULL
)

library(rrrSingleCellUtils)
library(Seurat)
library(ggplot2)
library(ggrepel)
library(tidyverse)
library(stringr)
library(harmony)
library(cowplot)

set.seed(888)

sysfonts::font_add("Arial Narrow", "ARIALN.TTF", bold = "ARIALNB.TTF")
showtext::showtext_auto()

theme_set(ggpubr::theme_pubr())

plot_cols <- c("#D43F3AFF", "#EEA236FF", "#357EBDFF", "#5CB85CFF", "#B8B8B8FF",
  "#9632B8FF",  "#46B8DAFF", "#90302DFF", "#A66D04FF", "#2D577FFF", "#3E7E3EFF",
  "#7D7D7DFF", "#6D1D87FF", "#097F9AFF", "#FF6E6AFF", "#FFBB70FF", "#68A4E3FF",
  "#79D379FF", "#CDCDCDFF", "#BF6BE2FF", "#69D1F3FF")

rDimPlot <- function(object, title = "Unlabeled Plot",
  label = T, pt.size = 1, ...) {
  if(length(levels(Idents(object))) < 22) {  
    p <- Seurat::DimPlot(
      object = object, 
      label = label, 
      pt.size = pt.size,
      cols = alpha(plot_cols, 0.6),
      ...) +
      ggtitle(title) +
      theme(legend.position = "none") +
      coord_fixed()
    print(p)
    return(p)
  } else {
    print("Too many identity classes to use this function. Requires <22.")
  }
}

rFeaturePlot <- function(object, features, title, pt.size = 1, order = T, ...) {
  p <- Seurat::FeaturePlot(
    object = object,
    features = features,
    pt.size = pt.size,
    order = order,
    cols = (c("lightgoldenrod", "darkred")),
    ...) +
    ggtitle(title) +
    coord_fixed()
  print(p)
  return(p)
}
```

# Isolate and subcluster the tumor cells

```{r, fig.width = 8, fig.width = 8}
# Isolate the tumor cells away from the comb Seurat object
selection <- readRDS("HuOsteo/selection.rds")
tumor <- subset(qs::qread("HuOsteo/comb.qs"), cells = selection)

# Normalize and scale tumor dataset
tumor <- tumor %>%
  NormalizeData() %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA(verbose = F)

# Run batch-correction/integration using harmony
tumor <- RunHarmony(tumor, group.by.vars = "src")
tumor <- tumor %>%
  RunUMAP(reduction = "harmony", dims = 1:20) %>%
  FindNeighbors(reduction = "harmony", dims = 1:20)

plot <- list()
opt <- tribble(
  ~res, ~score, ~clusters
)

for(i in 1:9) {
  opt[i, "res"] <- 0.1 * i
  tumor <- FindClusters(tumor, resolution = opt$res[i])
  plot[[i]] <- rDimPlot(tumor, paste("Resolution", opt$res[i]))
  opt[i, "score"] <- silhouette_mean(silhouette_seurat(tumor))
  opt[i, "clusters"] <- length(levels(tumor$seurat_clusters))
}

plot_grid(plotlist = plot, ncol = 3)
# Manually enforcing a resolution of 0.2 based on review of clusters
# and potential DEGs. Silhouette performs poorly for scoring intratumor
# heterogeneity.
tumor <- FindClusters(tumor, resolution = 0.2)

rDimPlot(tumor, "Subclustered tumor cells")

rDimPlot(tumor, "Subclustered tumor cells", split.by = "src", ncol = 4)

tumor <- kill_cc(tumor)
rDimPlot(tumor, "Sublustered tumor cells - cell cycle", group.by = "Phase")
```

# Generate gene set lists that identify each tumor cell cluster

NOTE: Since this is a sub-clustering of cells that are highly 
related, it would be especially appropriate here to validate our
sub-clustering with silhouette scoring or similar.

```{r}
Idents(tumor) <- tumor$seurat_clusters

if(!exists("tumor_marks.xlsx")) {
  tumor_marks <- FindAllMarkers(tumor)
  xlsx::write.xlsx(tumor_marks, "tumor_marks.xlsx")
}

tumor_marks <- xlsx::read.xlsx("tumor_marks.xlsx", 1)
tumor_marks <- as.tibble(
  tumor_marks[order(-tumor_marks$avg_log2FC),])
num_genes <- 20

for(i in levels(Idents(tumor))) {
  print(paste(
    "Top", num_genes, "UPREGULATED genes in cluster", i
  ))
  print(head(
    tumor_marks[tumor_marks$cluster == i,], n = num_genes
  ))
  print(paste(
    "Top", num_genes, "DOWNREGULATED genes in cluster", i
  ))
  print(tail(
    tumor_marks[tumor_marks$cluster == i,], n = num_genes
  ))
}
```

# From the cluster-marker gene lists, identify up- and down-regulated pathways

```{r}
dbs <- c("GO_Biological_Process_2021",
  "TF_Perturbations_Followed_by_Expression",
  "MSigDB_Hallmark_2020")

enriched <- list()
for (i in levels(Idents(tumor))) {
  for(d in 1:3) {
  degs <- tumor_marks$gene[tumor_marks$cluster == i]
  e <- enrichr(degs, dbs[d])
  enriched[[paste("Cluster", i, sep = "_")]] <- as_tibble(
    e[[1]])
  print(paste("Cluster", i, "top pathways", dbs[d]))
  print(as_tibble(e[[1]]), n = 20)
  }
}
```

# Compare results to those obtained using a pseudobulk procedure

First, generate DEG tables from pseudobulk clusters, treating all cells
from each cluster from each sample as independent groups of cells/data.
Then, graph each of those DEG datasets using a volcano plot.

```{r}
# Create pseudobulk tables
pseudobulk <- data.frame(gene = rownames(tumor@assays$RNA@counts))
rownames(pseudobulk) <- pseudobulk$gene

tumor$subcluster <- tumor$seurat_clusters
for(s in levels(as_factor(tumor$src))) {
  subtumor <- subset(tumor, subset = src == s)
  for(i in levels(subtumor$subcluster)) {
    if(length(which(subtumor$subcluster == i)) > 9) {
      subclust <- subset (subtumor, subcluster == i)
      lab <- paste0(s, "-C", i)
      pseudobulk[lab] <- rowSums(subclust@assays$RNA@counts)
    } else {
      print(paste("NOTE: Cluster", i, "from sample", s,
        "has less than 10 cells. This cluster will be skipped."))
    }
  }
}
pseudobulk <- pseudobulk[, -1]

coldata <- data.frame(sample = colnames(pseudobulk))
rownames(coldata) <- coldata$sample
coldata <- separate(coldata, "sample", c("sample", "condition"), sep = "-C")
coldata$type <- "single-read"
coldata <- coldata[, -1]

# Perform differential expression analysis between groups
ds2_data_sets <- list()
ds2_tests <- list()
ds2_results <- list()
p1 <- list()
for(i in unique(coldata$condition)) {
  coldata2 <- coldata
  coldata2["condition"][coldata2["condition"] == i] <- "target"
  coldata2["condition"][coldata2["condition"] != "target"] <- "other"

  ds2_data_sets[[i]] <- DESeq2::DESeqDataSetFromMatrix(
    countData = pseudobulk,
    colData = coldata2,
    design = ~ condition
  )

  ds2_tests[[i]] <- DESeq2::DESeq(ds2_data_sets[[i]])
  ds2_results[[i]] <- as.data.frame(DESeq2::results(ds2_tests[[i]]))

  ds2_results[[i]] <- ds2_results[[i]][complete.cases(ds2_results[[i]]), ]
  ds2_results[[i]] <- 
    ds2_results[[i]][order(-abs(ds2_results[[i]]$log2FoldChange)), ]
  lab <- head(ds2_results[[i]], n = 15)

  p1[[i]] <- ggplot(ds2_results[[i]], aes(log2FoldChange, -log10(padj))) +
    geom_point() +
    geom_label_repel(data = lab,
      aes(label = rownames(lab)),
      max.overlaps = 20) +
    ggtitle(paste("Cluster", i, "genes"))
  
  print(p1[[i]])
}
```

Identify pathways associated with changes in groups of those genes and plot results.

```{r}
degs <- list()

for(i in unique(coldata$condition)) {
  ds2_results[[i]] <- 
    ds2_results[[i]][order(-ds2_results[[i]]$log2FoldChange), ]
  degs[[i]] <- AnnotationDbi::select(org.Hs.eg.db::org.Hs.eg.db,
    keys = rownames(ds2_results[[i]]),
    columns = "ENTREZID",
    keytype = "ALIAS"
  )
  ds2_results[[i]]$ALIAS <- rownames(ds2_results[[i]])
  d <- inner_join(ds2_results[[i]], degs[[i]], by = "ALIAS")
  rownames(d) <- d$ENTREZID
  ds2_results[[i]] <- as.vector(d$log2FoldChange)
  names(ds2_results[[i]]) <- d$ENTREZID
  
  ego <- clusterProfiler::gseGO(
    geneList = ds2_results[[i]],
    OrgDb = org.Hs.eg.db::org.Hs.eg.db,
    ont = "BP",
    minGSSize = 100,
    maxGSSize = 500,
    pvalueCutoff = 0.05,
    verbose = FALSE
  )
  

}
```

