
## Test the correlation of expression within stromal cell types between groups
This is using pseudobulked data per stromal cell type per group as input. It is plotting a scatterplot of the genes between two groups for each stromal cell type to show how well they correlate across species/model and primary/met. It is also calculating fold change expression values for each comparison and using this as input for GSEA to show what pathways are differentially expressed between the two groups.

Mouse samples have their genes converted to human for this, and all samples are filtered to include only genes found across all datasets.
```{r group_corr_normal_subtypes}
cat_tib <- dplyr::tribble(
    ~category, ~subcategory,   ~cat_expl,
    "H",        "",            "Hallmark_paths",
    "C2",      "CP:KEGG",      "KEGG",
    "C3",      "TFT:GTRD",     "Transcription_factor_targets",
    "C5",      "GO:BP",        "GO Biological Process",
    "C5",      "GO:MF",        "GO Molecular Function",
    "C6",       "",            "Oncogenic_signature")

groups <-
    list.dirs(
        "output/seurat_objects/stromal_subtypes/",
        full.names = FALSE
    )[-1]

qs_file_list <-
    lapply(
        groups,
        function(group_name) {
            list.files(
                paste0(
                    "output/seurat_objects/stromal_subtypes/",
                    group_name,
                    "/"
                ),
                pattern = "qs",
                full.names = TRUE
            )
        }
    )
names(qs_file_list) <- groups

gsea_on_pseuobulked_sobjs(
    aggregation_column = "Ann_Level3",
    mouse_sample_regex = "mouse|^mm",
    qs_list = qs_file_list,
    parallel_threads = 4,
    gsea_categories = cat_tib,
    figure_dir = "output/figures/stromal_species_correlations/",
    corr_figure_name_stub = "celltype_correlation_",
    gsea_output = "output/stromal_species_correlations/gsea_results_normal.tsv"
)
```




## Test the correlation between within tumor subtypes between groups
This is using pseudobulked data per tumor subtype per group as input. It is plotting a scatterplot of the genes between two groups for each tumor subtype to show how well they correlate across species/model and primary/met. It is also calculating fold change expression values for each comparison and using this as input for GSEA to show what pathways are differentially expressed between the two groups.

Mouse samples have their genes converted to human for this, and all samples are filtered to include only genes found across all datasets.
```{r tumor_subtypes_corr}
base_dir <- "output/seurat_objects/final_tumor_vs_stroma/"
tumor_qs_files <-
    list(
        mm_prim = paste0(base_dir, "mm_prim_cancer_cells.qs"),
        mm_mets = paste0(base_dir, "mm_mets_cancer_cells.qs"),
        xeno_prim = paste0(base_dir, "xeno_prim_human.qs"),
        xeno_mets = paste0(base_dir, "xeno_mets_human.qs"),
        patient_prim = paste0(base_dir, "patient_prim_cancer_cells.qs"),
        patient_mets = paste0(base_dir, "patient_mets_cancer_cells.qs")
    )

gsea_on_pseuobulked_sobjs(
    aggregation_column = "Ann_Level3",
    mouse_sample_regex = "^mm",
    qs_list = tumor_qs_files,
    parallel_threads = 4,
    gsea_categories = cat_tib,
    figure_dir = "output/figures/tumor_subtypes_correlations/",
    corr_figure_name_stub = "tumor_subtypes_correlation_",
    gsea_output = "output/tumor_subtypes_correlations/gsea_results_tumor.tsv"
)
```
