
## Make a DimPlot with better default settings
```{r dimplot_better_function}
#| cache.vars = 'dimplot_better'
dimplot_better <- function(object,
                           group_by,
                           cols,
                           ncol = 1,
                           label.size = 2.5,
                           split_by = NULL,
                           ...) {
    DimPlot(object,
            group.by = group_by,
            split.by = split_by,
            ncol = ncol,
            label.size = label.size,
            label = T,
            repel = T,
            shuffle = T,
            label.box = T,
            cols = c(plot_cols, sample(rainbow(1000))),
            ...) +
        coord_fixed()
}
```

## Cell type annotaton data compilation
Human normal lung reference data downloaded from
Mouse lung reference data is from
https://pubmed.ncbi.nlm.nih.gov/33692365/
https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE151974
We kept only the normoxic samples for the analysis
```{r celltype_annotation_data}
#| cache.vars = c('human_lung_ref', 'mouse_lung_ref', 'human_tumor_cell', 'mouse_tumor_cell')
human_lung_ref <- qs::qread("input/downloads/normal_lung.qs")

human_lung_ref$free_annotation <-
    str_replace_all(human_lung_ref$free_annotation, c("/" = "_",
                                                  "\\+" = "_plus"))

human_lung_ref$free_annotation <-
    stringr::str_replace_all(
        human_lung_ref$free_annotation,
        c("^Alveolar Epithelial Type 1$" = "AlvEpithelial_T1",
          "^Alveolar Epithelial Type 2$" = "AlvEpithelial_T2",
          "^Signaling Alveolar Epithelial Type 2$" = "Signalling_AlvEpithelial",
          "^Basophil_Mast 1$" = "Basophil_Mast1",
          "^Basophil_Mast 2$" = "Basophil_Mast2",
          "^Bronchial Vessel 1$" = "Bronchial_Vessel1",
          "^Bronchial Vessel 2$" = "Bronchial_Vessel2",
          "^IGSF21_plus Dendritic$" = "Dendritic_cells",
          "^Myeloid Dendritic Type 1$" = "Myeloid_Dendritic1",
          "^Myeloid Dendritic Type 2$" = "Myeloid_Dendritic2",
          "^Natural Killer$" = "NK_cells",
          "^Natural Killer T$" = "NKT_cells",
          "^NK cells T$" = "NKT_cells",
          "^CD4_plus Memory/Effector T$" = "CD4T_Memory",
          "^CD4_plus Naive T$" = "CD4_Naive",
          "^CD8_plus Memory_Effector T$" = "CD8T_Memory",
          "^CD4_plus Memory_Effector T$" = "CD4T_Memory",
          "^CD8_plus Naive T$" = "CD8T_Naive",
          "^Proliferating NK_T$" = "Prolif_NKT",
          "^TREM2_plus Dendritic$" = "Dendritic_cells",
          "^Proliferating Macrophage$" = "Prolif_Macrophages",
          "^Proliferating Basal$" = "Prolif_Basal",
          "^Nonclassical Monocyte$" = "NonClassical_Monocytes",
          "^Classical Monocyte$" = "Classical_Monocytes",
          "^B$" = "B_cells",
          "^Plasma$" = "Plasma_Bcells",
          "^Vascular Smooth Muscle$" = "Vascular_SMC",
          "^Alveolar Fibroblast$" = "Alveolar_Fibroblasts",
          "^Capillary Aerocyte$" = "Capillary_Aerocytes",
          "^Plasmacytoid Dendritic$" = "Dendritic_cells",
          "^Signalling Alveolar Epithelial$" = "Epithelial_cells",
          "^Adventitial Fibroblast$" = "Fibroblasts",
          "^Signaling Alveolar_Epithelial$" = "Signalling_AlvEpithelial",
          "^Macrophage$" = "Macrophages",
          "^Differentiating Basal$" = "Differ_Basal",
          "^Airway Smooth Muscle$" = "Airway_SMC",
          "^Proximal Ciliated$" = "Proximal_Ciliated",
          "^Myofibroblast$" = "Myofibroblasts",
          "^Club$" = "Club_cells",
          "^Ionocyte$" = "Ionocyte",
          "^Ciliated$" = "Ciliated_cells",
          "^Pericyte$" = "Pericyte",
          "^Artery$" = "Artery_cells",
          "^Vein$" = "Vein_cells",
          "^Lymphatic$" = "Lymphatic_cells",
          "^Mucous$" = "Mucous_cells",
          "^Tcells$" = "T_cells",
          "^Capillary$" = "Capillary_cells"))

# #load in the human tumor cell from xenograft primary fro documentation
# sobj <- qs::qread("output/seurat_objects/comb_sobjs/xeno_prim_human.qs") %>%
#     subset(sample_name %in% c("S0170", "S0042"))

# #add a column with the cell type
# sobj$CellType <- "Tumor_cells"

# #save the object
# qs::qsave(sobj,
#           file = "input/downloads/human_tumor_fromxenograft.qs")

#load the object
human_tumor_cell <- qs::qread("input/downloads/human_tumor_fromxenograft.qs")

#mouse lung normal cells reference
mouse_lung_ref <- qs::qread("input/downloads/mouse_lung.qs")

mouse_lung_ref$CellType <-
            str_replace_all(mouse_lung_ref$CellType,
                            c("/" = "_",
                              "\\+" = "_plus",
                              "-" = "_"))

mouse_lung_ref$CellType <-
    str_replace_all(mouse_lung_ref$CellType,
                    c("^Col14a1_plus fibroblast$" = "Fibroblasts",
                      "^Col13a1_plus fibroblast$" = "Fibroblasts",
                      "^AT1$" = "AlvEpithelial_T1",
                      "^AT2 1$" = "AlvEpithelial_T2",
                      "^AT2 2$" = "AlvEpithelial_T2",
                      "^B cells$" = "B_cells",
                      "^Cap_a$" = "Capillary_Aerocytes", #capillary endothelial
                      "^CD4$" = "CD4T_cells1",
                      "^CD4 T cell 2$" = "CD4T_cells2",
                      "^CD8$" = "CD8T_cells1",
                      "^CD8 T cell 2$" = "CD8T_cells2",
                      "^DC1$" = "Dendritic_cells1",
                      "^DC2$" = "Dendritic_cells2",
                      "^gd T cell$" = "Tgd_cells",
                      "^Tgd$" = "Tgd_cells",
                      "^Neut 1$" = "Neutrophils1",
                      "^Neut 2$" = "Neutrophils2",
                      "^NK cells$" = "NK_cells",
                      "^Pericyte 1$" = "Pericyte1",
                      "^Pericyte 2$" = "Pericyte12",
                      "^Mast Ba2$" = "Basophil_Mast",
                      "^Mono$" = "Monocytes",
                      "^Alv Mf$" = "Alv_Macrophages",
                      "^Int Mf$" = "Int_Macrophages",
                      "^B cells$" = "B_cells",
                      "^AT$" = "Epithelial_cells",
                      "^Endothelial cells$" = "Endothelial_cells",
                      "^SMC$" = "Mural_cells",
                      "^ILC2$" = "Lymphoid_cells",
                      "^Dendritic cells$" = "Dendritic_cells",
                      "^AT$" = "EAlvEpithelial",
                      "^Cap$" = "Capillary_cells",
                      "^Lymph$" = "Lymph_cells",
                      "^Art$" = "Artery_cells",
                      "^Vein$" = "Vein_cells",
                      "^Myofibroblast$" = "Myo_Fibroblasts",
                      "^Club$" = "Club_cells",
                        "^Ionocyte$" = "Ionocyte",
                        "^Ciliated$" = "Ciliated_cells",
                        "^Pericytes$" = "Pericyte",
                        "^Mesothelial$" = "Mesothelial_cells",
                        "^Capillary$" = "Capillary_cells",
                        "^Mast_Basophils$" = "Basophil_Mast"))

# #make a reference for all tumor cells, using cell culture data documantation
# mouse_tumor_cell <-
#     qs::qread("output/seurat_objects/comb_sobjs/mm_prim.qs") %>%
#     subset(sample_name %in% c("S0200", "S0201"))

# #create a column with the cell type
# mouse_tumor_cell$CellType <- "Tumor_cells"

# #save the object
# qs::qsave(mouse_tumor_cell,
#           file = "input/downloads/mm_tumor_cellculture.qs")

#load the object S0200 and S0201 combined cell culture mouse tumor cells
mouse_tumor_cell <- qs::qread("input/downloads/mm_tumor_cellculture.qs")

```

## Function to annotate cell types with mouse and human lung
```{r annotate_function}
#| cache.vars = 'annotate_celltypes'
hpca <- celldex::HumanPrimaryCellAtlasData()
huim <- celldex::MonacoImmuneData()
ref4 <- SeuratObject::GetAssayData(human_tumor_cell)

hpca$label.main <-
            str_replace_all(hpca$label.main,
                            c("/" = "_",
                              "\\+" = "_plus",
                              "-" = "_"))
hpca$label.main <-
            stringr::str_replace_all(hpca$label.main,
                                     c("^T_cells$" = "T_cells",
                                       "^B_cell$" = "B_cells",
                                       "^NK_cell$" = "NK_cells",
                                       "^Monocyte$" = "Monocytes",
                                       "^DC$" = "Dendritic_cells",
                                       "^Pre_B_cell_CD34_$" = "CD34_minusB_cells",
                                       "^Pro_B_cell_CD34_plus$" = "CD34_plusB_cells",
                                       "^BM$" = "Bone_marrow",
                                       "^HSC_CD34_plus$" = "Hematopoietic_stem_cells",
                                       "^Pro_Myelocyte$" = "Myelocyte",
                                       "^MEP$" = "MEP_progenitors",
                                       "^GMP$" = "GMP_Progenitors",
                                       "^CMP$" = "CMP_Progenitors",
                                       "^BM & Prog.$" = "Bone_marrow",
                                       "^Macrophage$" = "Macrophages",
                                       "^Capillary$" = "Capillary_cells"))

huim$label.main <-
            str_replace_all(huim$label.main,
                            c("/" = "_",
                              "\\+" = "_plus",
                              "-" = "_"))
huim$label.main <-
    stringr::str_replace_all(huim$label.main,
                                c("^B cells$" = "B_cells",
                                "^Dendritic cells$" = "Dendritic_cells",
                                "^NK cells$" = "NK_cells",
                                "^T cells$" = "T_cells",
                                "^CD8_plus T cells$" = "CD8T_cells",
                                "^CD4_plus T cells$" = "CD4T_cells",
                                "^Progenitors$" = "Progenitor_cells",
                                "^Macrophage$" = "Macrophages",
                                "^Basophils$" = "Basophil_Mast",
                                "^Capillary$" = "Capillary_cells"))


mord <- celldex::MouseRNAseqData()
moim <- celldex::ImmGenData()
ref4 <- SeuratObject::GetAssayData(mouse_tumor_cell)
mord$label.main <-
    stringr::str_replace_all(mord$label.main,
                            c("^T cells$" = "T_cells",
                                "^B cells$" = "B_cells",
                                "^NK cells$" = "NK_cells",
                                "^Epithelial cells$" = "Epithelial_cells",
                                "^Dendritic cells$" = "Dendritic_cells",
                                "^Endothelial cells$" = "Endothelial_cells",
                                "^Progenitors$" = "Progenitor_cells",
                                "^Airway Smooth Muscle$" = "Airway_SMC",
                                "^Proximal Ciliated$" = "Ciliated_cells",
                                "^Myofibroblast$" = "Myo_Fibroblasts",
                                "^Club$" = "Club_cells",
                                "^Ionocyte$" = "Ionocyte",
                                "^Ciliated$" = "Ciliated_cells",
                                "^Microglia$" = "Macrophages",
                                "^Pericytes$" = "Pericytes",
                                "^Capillary$" = "Capillary_cells"))
moim$label.main <-
    stringr::str_replace_all(moim$label.main,
                            c("^B cells, pro$" = "B_cells",
                            "^DC$" = "Dendritic_cells",
                            "^NK cells$" = "NK_cells",
                            "^T cells$" = "T_cells",
                            "^Epithelial cells$" = "Epithelial_cells",
                            "^Endothelial cells$" = "Endothelial_cells",
                            "^B cells$" = "B_cells",
                            "^Stem cells$" = "Stem_cells",
                            "^Stromal cells$" = "Stromal_cells",
                            "^Mast cells$" = "Basophil_Mast",
                            "^NKT$" = "NKT_cells",
                            "^Tgd$" = "Tgd_cells",
                            "^ILC$" = "Lymphoid_cells",
                            "^Microglia$" = "Macrophages",
                            "^Capillary$" = "Capillary_cells"))

#function to annotate celltypes
annotate_celltypes <- function(sobject,
                               species = "",
                               with_tumor = FALSE,
                               ref,
                               labels,
                               aggr_ref = FALSE,
                               label_type = "label.main",
                               ...) {
    if (grepl("human", species)) {
        ref <- list(hpca,
                    huim)
        labels <- list(hpca[[label_type]],
                       huim[[label_type]])
        if (with_tumor == TRUE) {
            ref <- list(hpca,
                        huim,
                        SeuratObject::GetAssayData(human_tumor_cell))
            labels <- list(hpca[[label_type]],
                           huim[[label_type]],
                           human_tumor_cell$CellType)
        }
        if (species == "human_lung") {
            ref <- list(hpca,
                        huim,
                        SeuratObject::GetAssayData(human_lung_ref))
            labels <- list(hpca[[label_type]],
                           huim[[label_type]],
                           human_lung_ref$free_annotation)
        }
        if (species == "human_lung" && with_tumor == TRUE) {
            ref <- list(hpca,
                        huim,
                        SeuratObject::GetAssayData(human_tumor_cell),
                        SeuratObject::GetAssayData(human_lung_ref))
            labels <- list(hpca[[label_type]],
                           huim[[label_type]],
                           human_tumor_cell$CellType,
                           human_lung_ref$free_annotation)
        }
    } else if (grepl("mouse", species)) {
        ref <- list(mord,
                    moim)
        labels <- list(mord[[label_type]],
                       moim[[label_type]])
        if (with_tumor == TRUE) {
            ref <- list(mord,
                        moim,
                        SeuratObject::GetAssayData(mouse_tumor_cell))
            labels <- list(mord[[label_type]],
                           moim[[label_type]],
                           mouse_tumor_cell$CellType)
        }
        if (species == "mouse_lung") {
            ref <- list(mord,
                        moim,
                        SeuratObject::GetAssayData(mouse_lung_ref))
            labels <- list(mord[[label_type]],
                           moim[[label_type]],
                           mouse_lung_ref$CellType)
        }
        if (species == "mouse_lung" && with_tumor == TRUE) {
            ref <- list(mord,
                        moim,
                        SeuratObject::GetAssayData(mouse_tumor_cell),
                        SeuratObject::GetAssayData(mouse_lung_ref))
            labels <- list(mord[[label_type]],
                           moim[[label_type]],
                           mouse_tumor_cell$CellType,
                           mouse_lung_ref$CellType)
        }
    }
    annotation <-
        SingleR::SingleR(test = Seurat::as.SingleCellExperiment(sobject),
                         ref = ref,
                         labels = labels,
                         aggr.ref = aggr_ref,
                         ...)
    sobject$annotations <- annotation$labels
    sobject$cell_scores <-
        apply(X = annotation$scores,
              MARGIN = 1,
              function(x) max(x, na.rm = TRUE))
    return(sobject)
}
```

## RUN differential expression analysis on a Seurat object
```{r run_degs_function}
#| cache.vars: all_samples_csv, run_degs
all_samples_csv <- read_tsv("misc/allsample_details.txt",
                            show_col_types = FALSE)

run_degs <- function(sobject,
                     group_by = "") {

    pseudobulk <- sobject %>%
        Seurat::AggregateExpression(
            group.by = c("sample_name", "seurat_clusters"),
            slot = "counts",
            assay = "RNA"
        )
    pseudobulk <- pseudobulk[[1]]

    # run deseq2
    #for (cluster_no in unique(object$seurat_clusters)) {
    degs_output <-
        parallel::mclapply(unique(sobject$seurat_clusters),
                           mc.cores = parallelly::availableCores(),
                           mc.preschedule = FALSE,
                           function(cluster_no) {
        batch_df <- data.frame(sample_name = colnames(pseudobulk))

        batch_df_new <-
            tibble(col_label = colnames(pseudobulk),
                   sample_name = colnames(pseudobulk) %>%
                   str_remove("_[0-9]+$"),
                   cluster_no = str_remove(colnames(pseudobulk), ".+_")) %>%
            left_join(batch_df) %>%
            mutate(sample_name = str_replace_all(sample_name, "-", "_")) %>%
            mutate(organism = sobject$organism[1])

        #add the datasource and method as batch in batch_df_new from all_samples_csv for each sample_name
        new_all_samples_csv <- all_samples_csv %>%
            select(sample_name, method, data_source, model, location, organism)
        if (sobject$organism[1] == "mouse") {
            batch_df_new2 <- left_join(batch_df_new,
                                    new_all_samples_csv,
                                    by = c("sample_name", "organism")) %>%
                select(-sample_name, -method, -location, -data_source, -organism)
        } else {
            batch_df_new2 <- left_join(batch_df_new,
                                    new_all_samples_csv,
                                    by = c("sample_name", "organism")) %>%
                select(-sample_name, -method, -location, -model, -organism)
        }

        all(colnames(pseudobulk) == batch_df_new$col_label)

        batch_df_new2$group <- "other"
        batch_df_new2$group[batch_df_new2$cluster_no == cluster_no] <- "target"
        batch_df_new2$group <- factor(batch_df_new2$group)

        for (columns in colnames(batch_df_new2)) {
            if (unique(batch_df_new2[[columns]]) %>% length() == 1) {
                batch_df_new2[[columns]] <- NULL
            }
        }
        design <-
            as.formula(
                paste0(
                    "~",
                    paste(colnames(batch_df_new2)[c(-1, -2)], collapse = " + ")
                )
            )

        return(
            DESeq2::DESeqDataSetFromMatrix(countData = pseudobulk + 1,
                                           colData = batch_df_new2,
                                           design = design) %>%
            DESeq2::DESeq(quiet = TRUE) %>%
            DESeq2::results() %>%
            as.data.frame() %>%
            dplyr::arrange(dplyr::desc(log2FoldChange)) %>%
            tibble::rownames_to_column("gene") %>%
            na.omit()
        )
    })
    # subset and name the degs_output by the group_by variables
    names(degs_output) <- unique(sobject$seurat_clusters)

    return(degs_output)
}
```

## Function to run GSEA
```{r gsea_function}
#| cache.vars = 'run_gsea'
run_gsea <- function(degs_result,
                     category = "",
                     subcategory = "",
                     species = "") {
    gsea_ref <- msigdbr::msigdbr(species = species,
                                 category = category,
                                 subcategory = subcategory) %>%
        split(x = .$gene_symbol, f = .$gs_name)

    gsea_output <- list()
    #for (item in names(degs_result)) {
    gsea_output <-
        parallel::mclapply(names(degs_result),
                           mc.cores = parallelly::availableCores(),
                           mc.preschedule = FALSE,
                           function(item) {
            gsea_input <- as.vector(degs_result[[item]]$log2FoldChange)
            names(gsea_input) <- degs_result[[item]]$gene

            output <- fgsea::fgseaMultilevel(gsea_ref,
                                            gsea_input,
                                            minSize = 10,
                                            maxSize = 500,
                                            nPermSimple = 10000)
            gsea_output[[item]] <-
                output %>%
                dplyr::arrange(desc(NES)) %>%
                filter(padj < 0.05) %>%
                na.omit()
        })
    names(gsea_output) <- names(degs_result)

    return(gsea_output)
}
```

# GSEA Plots function for twoway dotplot and barplot
```{r}
#| cache.vars = c('gsea_dotplot', 'gsea_barplot')
gsea_dotplot <- function(data, x_col = "z_score") {
    lab4plot <-
        tibble(y = c(-2.5, 2.5),
               x = c(0.2, 0.2),
               label = c("Downregulated", "Upregulated"))

    plot_name <-
        ggplot() +
        geom_point(data = data,
                 aes(x = -1 * order,
                     y = get(x_col),
                     color = -log10(padj),
                     size = size),
                 stat = "identity",
                 alpha = 0.8) +
        coord_flip() +
        geom_text(data = data,                #text for pathways names
                    aes(x = -1 * order,
                        y = get(x_col) + (y_pos *5),
                        hjust = justify_y,
                        label = pathway),
                        size = 2,
                        fontface = "bold") +
        geom_text(data = lab4plot,    #text for upregulated/downregulated
                  aes(x = x,
                      y = y,
                      label = label),
                  fontface = "bold",
                  size = 2.5) +
        scale_fill_manual(values = plot_cols,
                          name = paste0(x_col, " > 0")) +
        theme(strip.background = element_rect(color = "white",
                                              fill = "white"),
              #legend.position = "none",
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.border = element_blank(),
              axis.line.x = element_line(color = "black"),
              axis.line.y = element_blank(),
              axis.title.x = element_text(size = 7),
              axis.ticks.y = element_blank(),
              plot.title = element_text(size = 7, face = "bold"),
              axis.text.y = element_blank(),
              axis.text.x = element_text(size = 7),
              legend.title = element_text(size = 7,
                                          face = "bold")) +
        labs(title = data$sample[1],
             y = "NES",
             x = "") +
        theme(plot.title = element_text(hjust = 0.5)) +
        ylim(-6, 6) +
        scale_color_gradient(low = plot_cols[2],
                             high =plot_cols[1])

    return(plot_name)
}


gsea_barplot <- function(data, x_col = "z_score") {
    lab4plot <-
        tibble(y = c(-3, 3),
               x = c(0.2, 0.2),
               label = c("Downregulated", "Upregulated"))

    plot_name <-
        ggplot() +
        geom_bar(data = data,
                 aes(x = -1 * order,
                     y = get(x_col),
                     fill = get(x_col) > 0),
                 stat = "identity",
                 alpha = 0.8) +
        coord_flip() +
        geom_hline(yintercept = 0,
                    color = "black",
                    linewidth = 0.5) +
        geom_text(data = data,                #text for pathways names
                    aes(x = -1 * order,
                        y = y_pos,
                        hjust = justify_y,
                        label = pathway),
                        size = 2,
                        fontface = "bold") +
        geom_text(data = lab4plot,    #text for upregulated/downregulated
                  aes(x = x,
                      y = y,
                      label = label),
                  fontface = "bold",
                  size = 2.5) +
        scale_fill_manual(values = plot_cols,
                          name = paste0(x_col, " > 0")) +
        theme(strip.background = element_rect(color = "white",
                                              fill = "white"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.border = element_blank(),
              axis.line.x = element_line(color = "black"),
              axis.line.y = element_blank(),
              axis.title.x = element_text(size = 7),
              axis.ticks.y = element_blank(),
              plot.title = element_text(size = 7, face = "bold"),
              axis.text.y = element_blank(),
              axis.text.x = element_text(size = 7),
              legend.title = element_text(size = 7,
                                          face = "bold")) +
        labs(title = data$sample[1],
             y = "NES",
             x = "") +
        theme(plot.title = element_text(hjust = 0.5)) +
        ylim(-6, 6)

    return(plot_name)
}
```

## Circle plot function
```{r circle_plot_function}
#| cache.vars: 'circle_plot'
circle_plot <- function(sobj,
                        cluster_column,
                        features) {
    avg_expr <-
        AverageExpression(
            sobj,
            assay = "RNA",
            layer = "scale.data",
            features = features,
            group.by = cluster_column
        ) %>%
        as.data.frame() %>%
        rownames_to_column("gene") %>%
        as_tibble() %>%
        pivot_longer(
            cols = -gene,
            names_to = "cluster",
            values_to = "expression"
        ) %>%
        mutate(cluster = str_replace(cluster, "RNA.g", "Cluster "))

    clusters <-
        sobj[[cluster_column]] %>%
        unlist() %>%
        unique() %>%
        as.character()

    perc_expressed <-
        lapply(clusters,
            function(x) {
                binarized <-
                    sobj[, sobj[[cluster_column]] == x] %>%
                    GetAssayData(slot = "data")
                binarized <- binarized[features, ]
                binarized <- binarized == 0
                perc_expressed <- 1 - rowSums(binarized) / ncol(binarized)
                return(
                    enframe(
                        perc_expressed,
                        name = "gene",
                        value = "perc_expressed"
                    ) %>%
                    mutate(cluster = paste0("Cluster ", x))
                )
            }) %>%
        bind_rows()

    ############### Use hclust to group genes and clusters
    # Calculate the distance matrix
    cluster_tree <-
        avg_expr %>%
        pivot_wider(
            names_from = gene,
            values_from = expression
        ) %>%
        column_to_rownames("cluster") %>%
        dist() %>%
        hclust()

    gene_tree <-
        avg_expr %>%
        pivot_wider(
            names_from = cluster,
            values_from = expression
        ) %>%
        column_to_rownames("gene") %>%
        dist() %>%
        hclust()

    left_join(avg_expr, perc_expressed, by = c("gene", "cluster")) %>%
        mutate(cluster = factor(cluster, levels = cluster_tree$label[cluster_tree$order]),
               gene = factor(gene, levels = gene_tree$label[gene_tree$order])) %>%
        ggplot(
            aes(
                x = cluster,
                y = gene,
                fill = expression
            )
        ) +
        geom_tile() +
        #coord_equal() +
        geom_point(aes(size = perc_expressed), shape = 1, color = "white") +
        scale_fill_gradient(
            low = "lightgray",
            high = "darkblue",
            name = "Scaled\nExpression"
        ) +
        scale_size_continuous(name = "Percent\nExpressed") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 10),
              axis.text.y = element_text(size = 10),
              legend.key = element_rect(fill = "lightgray")) +
        labs(x = NULL, y = NULL)
}
```

```{r panel_plot_function}
#| cache.vars: 'make_panel_plot'

#GSEA Miltilevel
cat_tib <- dplyr::tribble(
    ~category, ~subcategory,   ~cat_expl,
    "H",        "NA",          "Hallmark_paths",
    "C2",      "CP:KEGG",      "KEGG",
    "C3",      "TFT:GTRD",     "Transcription_factor_targets",
    "C6",       "NA",          "Oncogenic_signature")


make_panel_plot <- function(sobj, cluster_column, label, group) {
    volc_plots <- list()
    gsea_plots <- list()
    gsea_results <- list()
    heatmap_plots <- list()

    colors_use <-
        colorRampPalette(rev(RColorBrewer::brewer.pal(n = 7, name = "RdYlBu")))(100)

    colors_use <- c(colors_use[1:48], rep("gray", 4), colors_use[52:100])

    dimplot <- dimplot_better(sobj, group_by = cluster_column)
    dimplot_celltype <- dimplot_better(sobj, group_by = "new_annot_clust")

    degs_result <- run_degs(sobject = sobj, group_by = cluster_column)

    if (dir.exists(paste0("output/degs/", label))) {
        unlink(paste0("output/degs/", label), recursive = TRUE)
    }
    # Calculate the differentially expressed genes
    for (cluster_name in names(degs_result)) {
        if (!dir.exists(paste0("output/degs/", group, "/", label))) {
            dir.create(paste0("output/degs/", group, "/", label), recursive = TRUE)
        }
        write_tsv(
            degs_result[[cluster_name]],
            paste0("output/degs/", group, "/", label, "/", cluster_name, "_degs.tsv")
        )

        logfc_cutoff <- 0.6
        degs_result[[cluster_name]] <-
            degs_result[[cluster_name]] %>%
            mutate(
                diffexpressed = if_else(
                    log2FoldChange > logfc_cutoff & padj < 0.05,
                    "Upregulated",
                    if_else(
                        log2FoldChange < -1 * logfc_cutoff & padj < 0.05,
                        "Downregulated",
                        "Not Significant"
                    )
                )
            )

        #make the volcano plot
        volc_plots[[cluster_name]] <-
            ggplot(
                degs_result[[cluster_name]],
                aes(
                    x = log2FoldChange,
                    y = -log10(pvalue),
                    color = diffexpressed
                )
            ) +
            geom_vline(xintercept = c(-1 * logfc_cutoff, logfc_cutoff),
                      col = "gray",
                      linetype = "dashed") +
            geom_hline(yintercept = c(0.05),
                      col = "gray",
                      linetype = "dashed") +
            geom_point(size = 2) +
            scale_color_manual(values = c(plot_cols[1], "grey", plot_cols[2]),
                               name = "") +
            geom_text_repel(aes(label = gene))
    }

    marker_genes <-
        sapply(
            names(degs_result),
            simplify = FALSE,
            function(x) {
                x <- as.character(x)
                degs_result[[x]] %>%
                    mutate(cluster = paste0("Cluster ", x)) %>%
                    select(-baseMean, -lfcSE, -stat) %>%
                    filter(diffexpressed == "Upregulated") %>%
                    arrange(desc(log2FoldChange)) %>%
                    slice_head(n = 5)
        }) %>%
        bind_rows() %>%
        pull(gene)

    if (length(marker_genes) != 0) {
        circle_plot_out <-
            circle_plot(
                sobj,
                cluster_column = cluster_column,
                features = marker_genes)
    } else {
        circle_plot_out <- 
            ggplot(data = NULL) +
                   aes(x = as.factor(1), y = as.factor(1)) +
                   geom_text(aes(label = str_c("No marker genes for ", label)))
    }

    # GSEA analysis
    if (dir.exists(paste0("output/gsea/", label))) {
        unlink(paste0("output/gsea/", label), recursive = TRUE)
    }
    for (i in seq_len(nrow(cat_tib))){
        category <- cat_tib$category[i]
        subcategory <- cat_tib$subcategory[i]
        cat_expl <- cat_tib$cat_expl[i]
        if (subcategory == "NA") {
            gsea_result <-
                run_gsea(
                    degs_result = degs_result,
                    category = category,
                    species = sobj$organism[1]
                )
        } else {
            gsea_result <-
                run_gsea(
                    degs_result = degs_result,
                    category = category,
                    subcategory = subcategory,
                    species = sobj$organism[1]
                )
        }
        #test if directory to save the results exists
        if (!dir.exists(paste0("output/gsea/", group, "/", label))) {
            dir.create(paste0("output/gsea/", group, "/", label), recursive = TRUE)
        }

        gsea_result <-
            sapply(names(gsea_result),
                   simplify = FALSE,
                   function(x) {
                gsea_result[[x]] %>%
                    mutate(cluster = paste("Cluster", x),
                           pathway_group = cat_expl)
            })

        for (item in names(gsea_result)) {
              #save the results
            write_tsv(
                gsea_result[[item]],
                paste0(
                    "output/gsea/", group, "/", label, "/",
                    cat_expl, "_cluster_", item, "_gsea.tsv"
                )
            )
        }

        #########################################################
        # Drop in code to make heatmap here
        n_paths_per_cluster <- 20
        use_pathways <-
            gsea_result %>%
            bind_rows() %>%
            group_by(cluster) %>%
            mutate(abs_NES = abs(NES)) %>%
            slice_max(order_by = abs_NES, n = n_paths_per_cluster) %>%
            pull(pathway) %>%
            unique()

        max_abs_nes <-
            gsea_result %>%
            bind_rows() %>%
            pull(NES) %>%
            max() %>%
            ceiling()

        breaks_use <-
            seq(-max_abs_nes, max_abs_nes, length.out = 101)

        heatmap_plots[[cat_expl]] <-
            gsea_result %>%
            bind_rows() %>%
            filter(pathway %in% use_pathways) %>%
            select(
                -padj,
                -log2err,
                -ES,
                -size,
                -leadingEdge,
                -pval,
                -pathway_group) %>%
            pivot_wider(
                names_from = cluster,
                values_from = NES,
                values_fill = 0
            ) %>%
            column_to_rownames("pathway") %>%
            pheatmap::pheatmap(
                color = colors_use,
                breaks = breaks_use,
                display_numbers = TRUE,
                silent = TRUE,
                main = paste0(cat_expl),
                fontsize_row = 5,
                fontsize_number = 5
            )
    }

    panel_plots[[label]] <-
        patchwork::wrap_plots(
            dimplot,
            circle_plot_out,
            ggplotify::as.ggplot(heatmap_plots$Hallmark_paths),
            ggplotify::as.ggplot(heatmap_plots$KEGG),
            ggplotify::as.ggplot(heatmap_plots$Oncogenic_signature),
            ggplotify::as.ggplot(heatmap_plots$Transcription_factor_targets),
            ncol = 2) +
        patchwork::plot_annotation(title = label)
    ggsave(
        paste0("output/figures/gsea/", group, "/", label, "_six_panels.png"),
        width = 30,
        height = 50,
        limitsize = FALSE,
        plot = panel_plots[[label]])

    #save as qs
    qs::qsave(panel_plots[[label]],
              file = paste0("output/figures/gsea/", group, "/", label, "_six_panels.qs"))
}
```
