
This qmd will be used to read in the figures and perpare the html output to hand to our collaborators
## Before harmony
```{r, before_harmony_figure_panels}
#| cache.vars: c('preharmony_dimplot_list', 'featureplot1_tumor_list', 'featureplot2_immune_list')

sample_name_list <- c("patient_prim",
                      "patient_mets",
                      "mm_prim",
                      "mm_mets",
                      "xeno_prim_human",
                      "xeno_mets_human",
                      "xeno_prim_mouse",
                      "xeno_mets_mouse")

preharmony_dimplot_list <- list()
featureplot1_tumor_list <- list()
featureplot2_immune_list <- list()

for (group in sample_name_list) {
     preharmony_dimplot_list[[group]] <- 
        knitr::include_graphics(str_c("output/figures/combined_plots/",
                                    group,
                                    "/",
                                    "preHarmony_combined_plots.png"))

    featureplot1_tumor_list[[group]] <- 
        knitr::include_graphics(str_c("output/figures/combined_plots/",
                 group,
                 "/",
                 "preharmony_tumor_featureplots.png"))

    featureplot2_immune_list[[group]] <-
       knitr::include_graphics(str_c("output/figures/combined_plots/",
                 group,
                 "/",
                 "preharmony_immune_featureplots.png"))
}
```

### Pre Harmony Dimplots
```{r print_panel_plots1, results = 'asis'}
#| dependson: before_harmony_figure_panels
qreport::maketabs(preharmony_dimplot_list)
```

### Pre Harmony Tumor Featureplot
```{r, print_panel_plots2, results = 'asis'}
#| dependson: before_harmony_figure_panels
qreport::maketabs(featureplot1_tumor_list)
```

### Pre Harmony Immune Featureplot
```{r, print_panel_plots3, results = 'asis'}
#| dependson: before_harmony_figure_panels
qreport::maketabs(featureplot2_immune_list)
```

## After Harmony
```{r, after_harmony_figure_panels}
#| cache.vars: c('postharmony_dimplot_list', 'posth_featureplot1_tumor_list', 'posth_eatureplot2_immune_list')

postharmony_dimplot_list <- list()
posth_featureplot1_tumor_list <- list()
posth_eatureplot2_immune_list <- list()

for (group in c("patient_prim", "patient_mets", "mm_prim", "mm_mets")) {
    postharmony_dimplot_list[[group]] <- 
        knitr::include_graphics(str_c("output/figures/combined_plots/",
                group,
                "/",
                "combined_plots.png"))

    posth_featureplot1_tumor_list[[group]] <- 
        knitr::include_graphics(str_c("output/figures/combined_plots/",
                 group,
                 "/",
                 "tumor_featureplots.png"))

    posth_eatureplot2_immune_list[[group]] <-
        knitr::include_graphics(str_c("output/figures/combined_plots/",
                 group,
                 "/",
                 "immune_featureplots.png"))
}
```

### Post Harmony Dimplots
```{r print_panel_plots4, results = 'asis'}
#| dependson: after_harmony_figure_panels
qreport::maketabs(postharmony_dimplot_list)
```

### Post Harmony Tumor Featureplot
```{r, print_panel_plots5, results = 'asis'}
#| dependson: after_harmony_figure_panels
qreport::maketabs(posth_featureplot1_tumor_list)
```

### Post Harmony Immune Featureplot
```{r, print_panel_plots6, results = 'asis'}
#| dependson: after_harmony_figure_panels
qreport::maketabs(posth_eatureplot2_immune_list)
```


## Split of tumor and stroma plots
```{r, split_tumor_stroma}
#| cache.vars: c('all_dimplots_list', 'alltumor_featureplot_list', 'allimmune_featureplot_list')

all_dimplots_list <- list()
alltumor_featureplot_list <- list()
allimmune_featureplot_list <- list()

for (group in c("patient_prim", "patient_mets", "mm_prim", "mm_mets")) {
    for (cell_group_name in c("cancer_cells", "normal_cells")) {
        sub_group <- str_c(group, "_", cell_group_name)

        all_dimplots_list[[sub_group]] <- 
            knitr::include_graphics(str_c("output/figures/tumor_vs_stroma/",
                     group,
                     "/",
                     cell_group_name,
                     "_dimplot.png"))

        alltumor_featureplot_list[[sub_group]] <- 
            knitr::include_graphics(str_c("output/figures/tumor_vs_stroma/",
                     group,
                     "/",
                     cell_group_name,
                     "_tumor_featureplot.png"))

        allimmune_featureplot_list[[sub_group]] <-
            knitr::include_graphics(str_c("output/figures/tumor_vs_stroma/",
                     group,
                     "/",
                     cell_group_name,
                     "_immune_featureplot.png"))
    }
}


for (subgroup in c("xeno_prim_human",
                "xeno_prim_mouse",
                "xeno_mets_human",
                "xeno_mets_mouse")) {

    all_dimplots_list[[subgroup]] <-
        knitr::include_graphics(str_c("output/figures/tumor_vs_stroma/",
                 subgroup,
                 "/",
                 "dimplot.png"))

    alltumor_featureplot_list[[subgroup]] <-
        knitr::include_graphics(str_c("output/figures/tumor_vs_stroma/",
                 subgroup,
                 "/",
                 "tumor_featureplot.png"))

    allimmune_featureplot_list[[subgroup]] <-
        knitr::include_graphics(str_c("output/figures/tumor_vs_stroma/",
                 subgroup,
                 "/",
                 "immune_featureplot.png"))
}
```

### All Dimplots
```{r, print_panel_plots7, results = 'asis'}
#| dependson: split_tumor_stroma
qreport::maketabs(all_dimplots_list)
```

### All Tumor Featureplot
```{r, print_panel_plots8, results = 'asis'}
#| dependson: split_tumor_stroma
qreport::maketabs(alltumor_featureplot_list)
```

### All Immune Featureplot
```{r, print_panel_plots9, results = 'asis'}
#| dependson: split_tumor_stroma
qreport::maketabs(allimmune_featureplot_list)
```

## Clustree
```{r, clustree_plots}
#| cache.vars: clustrees
all_groups <-
    c("patient_prim_cancer_cells",
      "patient_mets_cancer_cells",
      "mm_prim_cancer_cells",
      "mm_mets_cancer_cells",
      "xeno_prim_human",
      "xeno_mets_human",
      "patient_prim_normal_cells",
      "patient_mets_normal_cells",
      "xeno_prim_mouse",
      "xeno_mets_mouse",
      "mm_prim_normal_cells",
      "mm_mets_normal_cells")

clustrees <- list()

for (group in all_groups) {
    clustrees[[group]] <-
        knitr::include_graphics(str_c("output/figures/tumor_vs_stroma/clustree/",
                        group,
                        ".png"))
}

```

### Clustree Plots
```{r, print_panel_plots10, results = 'asis'}
#| dependson: clustree_plots
qreport::maketabs(clustrees)
```

## Seurat Cluster Panel Plots
```{r, six_panel_plots}
#| cache.vars: sixpanel_plot_list
all_groups <-
    c("patient_prim_cancer_cells",
      "patient_mets_cancer_cells",
      "mm_prim_cancer_cells",
      "mm_mets_cancer_cells",
      "xeno_prim_human",
      "xeno_mets_human",
      "patient_prim_normal_cells",
      "patient_mets_normal_cells",
      "xeno_prim_mouse",
      "xeno_mets_mouse",
      "mm_prim_normal_cells",
      "mm_mets_normal_cells")

sixpanel_plot_list <- list()

for (group in all_groups) {
    sixpanel_plot_list[[group]] <-
        knitr::include_graphics(str_c("output/figures/gsea/",
                        group,
                        "/",
                        "seurat_clusters_six_panels.png"))
}

```

### Seurat Cluster Panel Plots
```{r, print_panel_plots11, results = 'asis'}
#| dependson: six_panel_plots
qreport::maketabs(sixpanel_plot_list)
```

## Clustering Normals AUCell and Correlation Plots
```{r, reclustering_normals_plots}
#| cache.vars: c('correlation_plot_list', 'Aucell_plot_list')
all_groups <-
    c("patient_prim_normal_cells",
      "patient_mets_normal_cells",
      "xeno_prim_mouse",
      "xeno_mets_mouse",
      "mm_prim_normal_cells",
      "mm_mets_normal_cells")

correlation_plot_list <- list()
Aucell_plot_list <- list()

for (group in all_groups) {
    path <- str_c("output/figures/relustering_normals/",
                  group)
    for (celltypes in c("Macrophages", "T_cells", "Fibroblasts", "Dendritic_cells")) {
        corr_file_path <- 
            str_c(path,
                  "/",
                  celltypes,
                  "_correlation.png")
        if (file.exists(corr_file_path)) {
            correlation_plot_list[[group]][[celltypes]] <-
                knitr::include_graphics(corr_file_path)
        }
        Aucell_file_path <- 
            str_c(path,
                  "/",
                  celltypes,
                  ".png")

        if (file.exists(Aucell_file_path)) {
            Aucell_plot_list[[group]][[celltypes]] <-
                knitr::include_graphics(Aucell_file_path)
        }
    }
}
```

### Correlation Plots: Patient primary Normal cells
```{r, print_panel_plots12, results = 'asis'}
#| dependson: reclustering_normals_plots
qreport::maketabs(correlation_plot_list[["patient_prim_normal_cells"]])
```

### AUCell Plots : Patient primary Normal cells
```{r, print_panel_plots13, results = 'asis'}
#| dependson: reclustering_normals_plots
qreport::maketabs(Aucell_plot_list[["patient_prim_normal_cells"]])
```

### Correlation Plots: Patient Mets Normal cells
```{r, print_panel_plots14, results = 'asis'}
#| dependson: reclustering_normals_plots
qreport::maketabs(correlation_plot_list[["patient_mets_normal_cells"]])
```

### AUCell Plots : Patient Mets Normal cells
```{r, print_panel_plots15, results = 'asis'}
#| dependson: reclustering_normals_plots
qreport::maketabs(Aucell_plot_list[["patient_mets_normal_cells"]])
```

### Correlation Plots: Xeno Primary Mouse
```{r, print_panel_plots16, results = 'asis'}
#| dependson: reclustering_normals_plots
qreport::maketabs(correlation_plot_list[["xeno_prim_mouse"]])
```

### AUCell Plots : Xeno Primary Mouse
```{r, print_panel_plots17, results = 'asis'}
#| dependson: reclustering_normals_plots
qreport::maketabs(Aucell_plot_list[["xeno_prim_mouse"]])
```

### Correlation Plots: Xeno Mets Mouse
```{r, print_panel_plots18, results = 'asis'}
#| dependson: reclustering_normals_plots
qreport::maketabs(correlation_plot_list[["xeno_mets_mouse"]])
```

### AUCell Plots : Xeno Mets Mouse
```{r, print_panel_plots19, results = 'asis'}
#| dependson: reclustering_normals_plots
qreport::maketabs(Aucell_plot_list[["xeno_mets_mouse"]])
```

### Correlation Plots: MM Primary Normal cells
```{r, print_panel_plots20, results = 'asis'}
#| dependson: reclustering_normals_plots
qreport::maketabs(correlation_plot_list[["mm_prim_normal_cells"]])
```

### AUCell Plots : MM Primary Normal cells
```{r, print_panel_plots21, results = 'asis'}
#| dependson: reclustering_normals_plots
qreport::maketabs(Aucell_plot_list[["mm_prim_normal_cells"]])
```

### Correlation Plots: MM Mets Normal cells
```{r, print_panel_plots22, results = 'asis'}
#| dependson: reclustering_normals_plots
qreport::maketabs(correlation_plot_list[["mm_mets_normal_cells"]])
```

### AUCell Plots : MM Mets Normal cells
```{r, print_panel_plots23, results = 'asis'}
#| dependson: reclustering_normals_plots
qreport::maketabs(Aucell_plot_list[["mm_mets_normal_cells"]])
```

## Normal Celltypes GSEA Panel Plots
```{r, celltype_panel_plots}
#| cache.vars: normalcell_gsea_panelplots

all_groups <-
    c("patient_prim_normal_cells",
      "patient_mets_normal_cells",
      "xeno_prim_mouse",
      "xeno_mets_mouse",
      "mm_prim_normal_cells",
      "mm_mets_normal_cells")

normalcell_gsea_panelplots <- list()

# the problem here is that I want to make a new plot list for each group and want to append the plots to it
for (group in all_groups) {
    for (celltypes in c("Macrophages", "T_cells", "Fibroblasts", "Dendritic_cells")) {
        file_path <- 
            str_c("output/figures/gsea/",
                  group,
                  "/",
                  celltypes,
                  "_six_panels.png")
        if (file.exists(file_path)) {
            normalcell_gsea_panelplots[[group]][[celltypes]] <-
                knitr::include_graphics(file_path)
        }
    }
}
```

### Patient Primary Normal Cells: gsea panel plots
```{r, print_panel_plots24, results = 'asis'}
#| dependson: celltype_panel_plots
qreport::maketabs(normalcell_gsea_panelplots[["patient_prim_normal_cells"]])
```

### Patient Mets Normal Cells: gsea panel plots
```{r, print_panel_plots25, results = 'asis'}
#| dependson: celltype_panel_plots
qreport::maketabs(normalcell_gsea_panelplots[["patient_mets_normal_cells"]])
```

### Xeno Primary Mouse: gsea panel plots
```{r, print_panel_plots26, results = 'asis'}
#| dependson: celltype_panel_plots
qreport::maketabs(normalcell_gsea_panelplots[["xeno_prim_mouse"]])
```

### Xeno Mets Mouse: gsea panel plots
```{r, print_panel_plots27, results = 'asis'}
#| dependson: celltype_panel_plots
qreport::maketabs(normalcell_gsea_panelplots[["xeno_mets_mouse"]])
```

### MM Primary Normal Cells: gsea panel plots
```{r, print_panel_plots28, results = 'asis'}
#| dependson: celltype_panel_plots
qreport::maketabs(normalcell_gsea_panelplots[["mm_prim_normal_cells"]])
```

### MM Mets Normal Cells: gsea panel plots
```{r, print_panel_plots29, results = 'asis'}
#| dependson: celltype_panel_plots
qreport::maketabs(normalcell_gsea_panelplots[["mm_mets_normal_cells"]])
```


